A fast unified model for parsing and sentence understanding

### 2016-03-15 writing/hard_stack_paper/paper.tex
Typing up Abi's comments.

Tree-structured neural networks exploit valuable syntactic parse information as they interpret the meanings of sentences. However, they suffer two key technical problems: they can only operate on parsed sentences and do not directly support batched computation, making them slow and unwieldy for large-scale NLP tasks. We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single model by integrating tree-structured sentence interpretation into the linear sequential structure of a transition-based parser. This model  supports batched computation, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy. We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence encoding models.

@@

A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \citep{socher2011semi,sutskever2014sequence,vinyals2015neural}. This component, the sentence encoder, is generally formulated as a learned parametric function from word vectors to sentence vectors, and this function can take a range of different forms. Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory (LSTM RNNs, \citet{hochreiter1997long}, see Figure~\ref{fig:batching:good}), which accumulate information over the sentence sequentially, convolutional neural networks over words or characters (\citet{kalchbrenner2014convolutional}, \todo{[SB] Cite LeCun CharCNN}), which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks (TreeRNNs \citet{goller1996learning,socher2011semi}, see Figure~\ref{fig:batching:bad}), which propagate information up a binary parse tree structure over words. Of these, it is easy to argue that the TreeRNN is the principled choice---meaning in natural language sentences is widely believed to be constructed incrementally according to a tree structure \citep{Partee84,Janssen97}, and so a model which instantiates that tree structure should have a maximally easy time learning to construct meaning.

@@

A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and transitions $\mathbf t = (t_1, \dots, t_T)$. It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure. For any binary-branching tree structure over $N$ words, this requires $T = 2N - 1$ transitions.

@@

The core SPINN model implements a sequential shift-reduce parser which operates over sentences. It is designed to produce a vector representation of the sentence as output,as opposed to a tree as in standard shift-reduce \textit{parsing}. This relies on two key modifications to the shift-reduce algorithm:
\begin{itemize}
\item The parser's intermediate representations, including the representations of partial tree structures on the stack, are fixed-length vectors.
\item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a neural network.
\end{itemize}

@@

The buffer is populated with the words in each sentence before the first step of computation. We use word representations from the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors that are stored side-by-side in the buffer and used as inputs to a TreeLSTM composition function:

@@

In the SPINN, each row of the stack and each row of the buffer contains a pair of vectors $\langle \vec{h}, \vec{c}\rangle$, which together represent some word or nonterminal node in the parse tree of the sentence. \todo{[SB] Clarify that this is needed because of the TreeLSTM fn.}

@@

This paper evaluates the SPINN within the context of \textit{natural language inference} (also known as \textit{recognizing textual entailment}), a sentence pair classification task. To classify a sentence pair, a feature vector is first constructed. This feature vector is based on the final representations of each of the two sentences---the representations that appear at the head of the stack for each sentence after the final transition. In particular, the $\vec{h}$ portions of these representations are used. The final feature vector consists of the concatenation of these two vectors, $\vec{x}_{\text{premise}}$ and $\vec{x}_{\text{hypothesis}}$, their difference, and their elementwise product:

\begin{equation}
\vec{x}_{classifier} = 
\colvec{4}
    {\vec{h}_{\text{premise}}}
    {\vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} - \vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} \odot \vec{h}_{\text{hypothesis}}}
\end{equation}

@@

@@

The model described so far is the SPINN-NP model, which uses sequences of \shift~and \reduce~operations from an external parser. To build the fully independent SPINN model, we simply add a single-layer neural network classifier that chooses which operation to perform at each step using the state of the tracking LSTM:
\begin{equation}
\vec{p}_{trans} = \text{softmax}(W_{trans}\vec{h}_{track})
\end{equation}

The resulting vector $\vec{p}_{trans}$ is a probability distribution over parsing decisions (i.e., predicting whether to \shift~or \reduce~at the current timestep). In the full SPINN model, we follow whichever transition has a higher predicted probability in this distribution. This prediction module is trained to mimic the decisions of an external parser. In particular, we use the transition sequences corresponding to the constituency parses included in the SNLI corpus. The classifier is trained using a cross-entropy objective:
\begin{equation}
  L_{trans} = - \log \vec{p}_{trans}[t^*]
\end{equation}
where $t^*$ is the correct transition specified by the gold parse, and $\vec{p}[\cdot]$ represents vector indexing.

@@

\paragraph{The size of the stack}
The size of the stack should be $N$ for sentences of $N$ words, in case the first \reduce~operation merges the final two words. The size of the buffer should be $N$.

\paragraph{Converting parses to transition sequences}

@@

A naive implementation of the SPINN model would require simulating a stack of size $N$ for each input sentence. In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations. This implies a total space requirement of $M \times N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$. Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack).

@@

In its bare form, before the addition of the tracking LSTM, the SPINN (or, SPINN-NP-NT, for \textit{no parsing, no tracking}) computes the precisely the same functions as a conventional tree-structured neural network model, and has the same learning dynamics: the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function). While it adds no expressive power on its own, this design improves upon the standard TreeRNN implementation in three ways:

@@

\subsection{Related work: Transition-based parsing and neural networks.}

@@

\subsection{Natural language inference and SNLI}

@@

We evaluate four models on the SNLI task. Each uses 300d hidden states:
\begin{itemize}
\item A baseline LSTM model (similar to that of \citet{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM-RNN sequence model.
\item The minimal SPINN-NP-NT (\textit{no parsing, no tracking}) model, which uses transitions from an external parser, and has no tracking LSTM to inform semantic composition. This model is equivalent to a TreeLSTM \citep{tai2015improved}.
\item The SPINN-NP model, which includes a tracking LSTM which guides composition, making it a hybrid tree-sequence model.
\item The full SPINN model, which makes its own parsing decisions, and does not depend on an external parser at test time, making it slightly weaker but a practical choice in applied settings.
\end{itemize}

We compare our models against a range of baselines, including the strongest published non-neural network-based result from \citep{snli:emnlp2015}, a set of previous neural network models built around sentence encoders, and a set of models built around soft attention architectures. We should note that these attention-based models are by far the strongest published on SNLI to date, but we do not wish to compare our model with those models directly. SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks. The attentional models are narrowly adapted to the sentence-pair structure of SNLI examples and cannot act as such general sentence encoders. We focus on models which do support general sentence encoding in our analysis.

\paragraph{Results} Table~\ref{tab:results} shows our results. Our primary evaluation metric for all models is three-way classification accuracy on the SNLI test set. The full SPINN model also generates parsing predictions, and we report a measure of agreement between this model's predictions and the Stanford Parser's predictions on the same data (calculated as classification accuracy over \shift~and \reduce~operations). We also report the number of trained parameters in each model and training set accuracy.

@@

At batch sizes of interest for test-time applications (e.g., 512 or 1024 examples), we observe a 10--20x speedup relative to the standard CPU implementation. Figure~\ref{speed} shows a substantial difference in runtime scaling as batch sizes increase which is typical of GPU computation. We hope that this improved batchable TreeRNN algorithm will enable further rapid research in tree-structured neural networks.

@@

\paragraph{Future work} Because our aim for this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention, despite its demonstrated effectiveness of the SNLI task. However, the choice of whether to use soft attention and the choice of what representations to perform attention over are orthogonal, and we expect that it should be possible to productively combine our model with soft attention to yield state-of-the-art performance.

@@

>>>>

Tree-structured neural networks exploit valuable syntactic parse information as they interpret the meanings of sentences. However, they suffer from two key technical problems: they can only operate on parsed sentences and do not directly support batched computation, making them slow and unwieldy for large-scale NLP tasks. We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single model by integrating tree-structured sentence interpretation into the linear sequential structure of a transition-based parser. This model  supports batched computation, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy. We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence encoding models.

@@

A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \citep{socher2011semi,sutskever2014sequence,vinyals2015neural}. This component, the sentence encoder, is generally formulated as a learned parametric function from a sequence of word vector to a sentence vector, and this function can take a range of different forms. Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory (LSTM RNNs, \citet{hochreiter1997long}, see Figure~\ref{fig:batching:good}), which accumulate information over the sentence sequentially, convolutional neural networks over words or characters (\citet{kalchbrenner2014convolutional}, \todo{[SB] Cite LeCun CharCNN}), which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks (TreeRNNs \citet{goller1996learning,socher2011semi}, see Figure~\ref{fig:batching:bad}), which propagate information up a binary parse tree structure over words. Of these, it is easy to argue that the TreeRNN is the principled choice---meaning in natural language sentences is widely believed to be constructed incrementally according to a tree structure \citep{Partee84,Janssen97}, and so a model which instantiates that tree structure should have a maximally easy time learning to construct meaning.

@@

A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and transitions $\mathbf t = (t_1, \dots, t_T)$. It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure. For any binary-branching tree structure over $N$ words, this requires $T = 2N - 1$ transitions. \todo{[JG] Try to introduce transitions before referring to the transition var $\mathbf t$. Maybe say that each $t\in\{\shift, \reduce\}$ }

@@

The core SPINN model implements a sequential shift-reduce parser which operates over sentences. It is designed to produce a vector representation of the sentence as output, as opposed to a tree as in standard shift-reduce parsing. This relies on two key modifications to the shift-reduce algorithm:
\begin{itemize}
\item The parser's intermediate representations, including the representations of partial tree structures on the stack, are fixed-length vectors.
\item The \reduce~operation combines two such representations of trees on the stack into a larger tree using a neural network.
\end{itemize}

@@

The buffer is populated with the words in each sentence before the first step of computation. We use word representations from the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to a TreeLSTM composition function:

@@

In the SPINN, each row of the stack and each row of the buffer contains a pair $\langle \vec{h}, \vec{c}\rangle$, which together represent some word or nonterminal node in the parse tree of the sentence. \todo{[SB] Clarify that this is needed because of the TreeLSTM fn.}

@@

This paper evaluates the SPINN within the context of \textit{natural language inference} (also known as \textit{recognizing textual entailment}), a sentence pair classification task. To classify a sentence pair, a feature vector $\vec{x}_{\text{classifier}}$ is first constructed. This feature vector is based on the final representations of each of the two sentences---the representations that appear at the head of the stack for each sentence after the final transition. In particular, the $\vec{h}$ portions of these representations are used. The final feature vector consists of the concatenation of these two vectors, $\vec{h}_{\text{premise}}$ and $\vec{h}_{\text{hypothesis}}$, their difference, and their elementwise product:

\begin{equation}
\vec{x}_{\text{classifier}} = 
\colvec{4}
    {\vec{h}_{\text{premise}}}
    {\vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} - \vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} \odot \vec{h}_{\text{hypothesis}}}
\end{equation}

@@

\todo{[SB] Good point to reference fig 2a?}

@@

The model described so far is the SPINN-NP model, which uses sequences of \shift~and \reduce~operations from an external parser. To build the fully independent SPINN model, we simply add a single-layer neural network classifier that chooses which operation to perform at each step using the state of the tracking LSTM:
\begin{equation}
\vec{p}_{\text{trans}} = \text{softmax}(W_{\text{trans}}\vec{h}_{\text{track}})
\end{equation}

The resulting vector $\vec{p}_{\text{trans}}$ is a probability distribution over parsing decisions (i.e., predicting whether to \shift~or \reduce~at the current timestep). In the full SPINN model, we follow whichever transition has a higher predicted probability in this distribution. This prediction module is trained to mimic the decisions of an external parser. In particular, we use the transition sequences corresponding to the constituency parses included in the SNLI corpus. The classifier is trained using a cross-entropy objective:
\begin{equation}
  L_{\text{trans}} = - \log \vec{p}_{\text{trans}}[t^*]
\end{equation}
where $t^*$ is the correct transition specified by the gold parse, and $\vec{p}[\cdot]$ represents vector indexing.

@@

\paragraph{The size of the stack}
The size of the stack should be $N$ for sentences of $N$ words, in case the first \reduce~operation merges the final two words. The size of the buffer should be $N$. \todo{[SB] Height, not storage size.}

\paragraph{Converting parses to transition sequences}
\todo{[SB] Remove all Model N references.}

@@

A naive implementation of the SPINN model would require simulating a stack of size $N$ for each input sentence. In order to update the model weights via backpropagation, we would also need to maintain all intermediate stack representations for later chain rule calculations. This implies a total space requirement of $M \times N \times T \times D$, which is prohibitively large for significant batch sizes $M$ or sentence lengths $N$. Such a naive implementation would also require duplicating a largely unchanged stack at each timestep (as \shift~and \reduce~operations only affect the top of the stack). \todo{[JG] Remind reader what T and D are here.}

@@

\todo{[SB] Make equivalence claim more explicit. Clean up bullets.}

In its bare form, before the addition of the tracking LSTM, the SPINN (or, SPINN-NP-NT, for \textit{no parsing, no tracking}) computes precisely the same functions as a conventional tree-structured neural network model, and has the same learning dynamics: the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function). While it adds no expressive power on its own, this design improves upon the standard TreeRNN implementation in three ways:

@@

\subsection{Related work: Transition-based parsing and neural networks.}
\todo{[SB] Move!}

@@

\subsection{Natural language inference and SNLI}
\todo{[SB] Move up.}

@@

We evaluate four models on the SNLI task. Each uses 300d hidden states: \todo{[SB] d or D?}
\begin{itemize}
\item A baseline LSTM model (similar to that of \citet{snli:emnlp2015}) that uses the same classifier architecture as our models, but encodes sentences using a single layer LSTM-RNN sequence model.
\item The minimal SPINN-NP-NT (\textit{no parsing, no tracking}) model, which uses transitions from an external parser, and has no tracking LSTM to inform semantic composition. This model is equivalent to a TreeLSTM \citep{tai2015improved}.
\item The SPINN-NP model, which includes a tracking LSTM which guides composition, making it a hybrid tree-sequence model.
\item The full SPINN model, which makes its own parsing decisions, and does not depend on an external parser at test time, making it slightly weaker but a practical choice in applied settings.
\end{itemize}

We compare our models against a range of baselines, including the strongest published non-neural network-based result from \citep{snli:emnlp2015}, a set of previous neural network models built around sentence encoders, and a set of models built around soft attention architectures. We should note that these attention-based models are by far the strongest published on SNLI to date, but we do not wish to compare our model with those models directly. \todo{[SB] But we do. Rephrase.} SNLI was originally presented as an evaluation and development task for sentence encoding models, on the premise that sentence encoding models that do well on SNLI capture sentence meaning in a general sense, and are well suited to use on a range of NLP tasks. The attentional models are narrowly adapted to the sentence-pair structure of SNLI examples and cannot act as such general sentence encoders. We focus on models which do support general sentence encoding in our analysis.

\paragraph{Results} Table~\ref{tab:results} shows our results. Our primary evaluation metric for all models is three-way classification accuracy on the SNLI test set. The full SPINN model also generates parsing predictions, and we report a measure of agreement \todo{[SB] Clarify.} between this model's predictions and the Stanford Parser's predictions on the same data (calculated as classification accuracy over \shift~and \reduce~operations). We also report the number of trained parameters in each model and training set accuracy.

@@

At batch sizes of interest for test-time applications (e.g., 512 or 1024 examples), we observe a 10--20x speedup relative to the standard CPU implementation. Figure~\ref{fig:speed} shows a substantial difference in runtime scaling as batch sizes increase which is typical of GPU computation. We hope that this improved batchable TreeRNN algorithm will enable further rapid research in tree-structured neural networks.

@@

\paragraph{Future work} Because this paper aims to introduce a general purpose model for sentence encoding, we did not pursue the use of soft attention, despite its demonstrated effectiveness of the SNLI task. However, the choice of whether to use soft attention and the choice of what representations to perform attention over are orthogonal \todo{[SB] Rephrase prev clause.}, and we expect that it should be possible to productively combine our model with soft attention to yield state-of-the-art performance.

@@

\todo{[SB] Fix run-ons here.}
